Vowel Elision in Yoruba Collocations

### 2023-02-26 paper.tex
Add newest changes

\section{Computing Ideal Surface Forms}

@@

@@

\textbf{Change}, to some extent, is functionally a composition of the previous
two operations, replacing one segment with another one. However, this does not
reflect how such transitions are modeled in the literature, using the
categorization of segments into their features. To accommodate this without
undertaking the modeling of autosegmental phonology
\cite{goldsmith1990autosegmental} in a computationally tractable way, we
manually partition the set of segments into phonologically similar (not
necessarily disjoint) sets, and attempt to map largely within these sets.
Illustrating concretely, we might classify all vowels in the language into one
set, and all the sonorants into another, so our two sets would look something
like this:

>>>>

\section{Computing Optimal Surface Forms}

@@

\subsubsection{Core Operations for Generating Candidates}

@@

\textbf{Change}, to some extent, is functionally a composition of the previous
two operations, replacing one segment with another one. However, this does not
reflect how such transitions are modeled in the literature, using the
categorization of segments into their features. To accommodate this without
undertaking the modeling of autosegmental phonology
\cite{goldsmith1990autosegmental} in a computationally tractable way, we
manually partition the set of segments into phonologically similar (not
necessarily disjoint) sets, and attempt to map largely within these sets.
Illustrating concretely, we might classify all vowels in the language into one
set, and all sonorants into another:

\begin{enumerate}
    \item \textipa{a o O i E u}
    \item \textipa{l m n w j}
\end{enumerate}

The probability of a segment changing internally to either of these sets is
high, but the probability of a segment changing across one or the other is much
lower. We note that it may be necessary to adjust the weighting of these
probabilities based on the language, or on cross-linguistic data more generally.

\textbf{Metathesize}, finally, can similarly be viewed as a composition of the
delete and epenthesize, or as a special case of the change operation. It
differs, of course, in its ability to make targeted transpositions which are
thought, based on cross-linguistic data, to be atomic (meaning indivisible, or
inseparable into components) operations of their own. In particular, it's
significantly more likely to swap two adjacent segments, again deploying
cross-linguistic observations.

\subsubsection{Generating Candidates}

With these operations in hand, we're now ready to address the interesting
portion: generating candidates and identifying the ideal one. First up, a short
description of how candidates are generated. We start with the underlying
representation. We then use each of \texttt{delete}, \texttt{epenthesize},
\texttt{change}, and \texttt{metathesize} on the underlying form, 4 times each.
This produces an initial candidate set of 16 candidates. We repeat the process
again on each entry, producing 256 entries. We repeat two more times, yielding
65536 total candidates. We're now ready to evaluate this candidate set against
our ranked constraints, but before we do so, it is appropriate to discuss how,
precisely, we represent constraints in a computable way.

\subsubsection{Encoding Constraints}

We represent constraints as functions. Specifically, these functions take as
input a candidate surface form, compare it against the underlying
representation, and return a count of violations. The precise logic used to
compare a candidate surface form against the underlying representation will
depend on the nature of each constraint. For example, for the constraint
\textsc{Dep}, the function would compare the underlying form to the candidate
surface form and determine if the latter has any segments which are not present
in the former. Of course, it must also determine whether the seemingly
epenthetic segment is the result of \texttt{change} or \texttt{metathesize},
and if so, avoid outputting a violation mark for that segment. This could be
determined by comparing segments around the errant one for equality. If
corresponding segements are equal at corresponding indices of the underlying
representation and the candidate surface form, modulo the errant segment, we
can be reasonably sure that \texttt{change} has taken place, otherwise, if
they are not, we can be reasonably sure that \texttt{metathesize} is the
culprit.

\subsubsection{Evaluating Candidates}

We are now in the appropriate setting to identify the correct output form.
In order to do this, we must evaluate our entire set of constraints against
all of the 65536 candidate surface forms we generated previously. This will
generate a sequence of constraint violations, one for each candidate surface
form.
